---
title: "Doubly robust etc"
output: html_notebook
---

Implementation of methods from the paper Chen, Y., Li, P., & Wu, C. (2019). Doubly Robust Inference With Nonprobability Survey Samples. Journal of the American Statistical Association, 1-11.


# Packages

```{r}
library(maxLik)
library(rootSolve)
library(sampling)
library(compiler)
```



# Simulated data

## Population data

```{r}
set.seed(666)
N <- 20000
## this may vary but not here
n_a <- 1000
n_b <- 1000
### 
z1 <- rbinom(N,1,0.7)
z2 <- runif(N,0,2)
z3 <- rexp(N,1)
z4 <- rchisq(N, 4)
x1 <- z1 
x2 <- z2 + 0.3*x1
x3 <- z3 + 0.2*(x1 + x2)
x4 <- z4 + 0.1*(x1 + x2 + x3)
epsilon <- rnorm(N)
## controlled at 0.3 (10.4), 0.5 (5.4) and 0.8 (2.4) for the simulation studies
sig30 <- 10.4
sig50 <- 5.4
sig80 <- 2.4
y30 <- 2 + x1 + x2 + x3 + x4 + sig30*epsilon
y50 <- 2 + x1 + x2 + x3 + x4 + sig50*epsilon
y80 <- 2 + x1 + x2 + x3 + x4 + sig80*epsilon

sim_data <- data.frame(y30, y50, y80, x1, x2, x3, x4, z1, z2, z3, z4)
head(sim_data)
```

## Non-probability sample

```{r}
set.seed(1)
eta_a <- -4.406 + 0.1*x1 + 0.2*x2 + 0.1*x3 + 0.2*x4
prob_a <- exp(eta_a) / (1 + exp(eta_a))
sample_a_flag <- UPpoisson(prob_a)
sample_a <- sim_data[sample_a_flag == 1,]
sample_a$prob <- prob_a[sample_a_flag == 1]
sample_a$weight <- 1/sample_a$prob
sample_a$nonprob <- 1
sample_a
```

## Probability sample

```{r}
set.seed(1)
z_b <- x3  + 0.215
prob_b <- inclusionprobabilities(z_b,n_b)
sample_b_flag <- UPsystematic(prob_b)  
sample_b <- sim_data[sample_b_flag == 1,]
sample_b$prob <- prob_b[sample_b_flag == 1]
sample_b$weight <- 1/sample_b$prob
sample_b$nonprob <- 0
sample_b
```

Comparison

```{r}
c(true = mean(y30), nonprob  = mean(sample_a$y30), prob = weighted.mean(sample_b$y30,sample_b$weight))
```

# Propensity score method

Just to see if it works

```{r}
samples <- rbind(sample_a,sample_b)
samples$x0 <- 1
```

Score function using 1, x1, x2, x3


```{r}
pseudo_ll_propensity <- function(x) {
 
  ## totals for non-prob sample
  totals_a <- colSums(samples[samples$nonprob == 1, c("x0", "x1", "x2", "x3", "x4")])
  
  ## Xs for probability samples
  X_b <- samples[samples$nonprob == 0, c("x0", "x1", "x2", "x3", "x4")]
  weight_b <- samples[samples$nonprob == 0, c("weight")]
  probs <- exp(x %*% t(X_b)) / (1 + exp(x %*% t(X_b)))
  totals_b <- colSums(weight_b * probs * X_b)
  totals_a - totals_b
} 

res <- multiroot(f = pseudo_ll_propensity, start = rep(0,5))
res
```


```{r}
X_a <- samples[samples$nonprob == 1, c("x0", "x1", "x2", "x3", "x4")]
probs_a <- as.numeric(exp(as.matrix(X_a)  %*% res$root) / ( 1 + exp(as.matrix(X_a)  %*% res$root)))
weighted.mean(samples[samples$nonprob == 1, c("y30")], 1/probs_a)
```

## Calibration approach

```{r}
pseudo_ll_calib <- function(par) {
  X_a <- samples[samples$nonprob == 1, c("x0", "x1", "x2", "x3", "x4")]
  totals <- colSums(cbind(1,sim_data[,c("x1", "x2", "x3", "x4")]))
  probs <- exp(par %*% t(X_a)) / (1 + exp(par %*% t(X_a)))
  colSums(X_a/probs) - totals
}

res2 <- multiroot(f = pseudo_ll_calib, start = rep(0,5))
res2

par <- res2$root
X_a <- samples[samples$nonprob == 1, c("x0", "x1", "x2", "x3", "x4")]
probs <- as.numeric(exp(par %*% t(X_a)) / (1 + exp(par %*% t(X_a))))
colSums(X_a/probs)
weighted.mean(samples[samples$nonprob == 1, "y30"], as.numeric(1/probs))


```

Add misclass error


```{r}
samples$x1er <- samples$x1 == 0

pseudo_ll_calib <- function(par) {
  X_a <- samples[samples$nonprob == 1, c("x0", "x1er", "x2", "x3", "x4")]
  totals <- colSums(cbind(1,sim_data[,c("x1", "x2", "x3", "x4")]))
  probs <- exp(par %*% t(X_a)) / (1 + exp(par %*% t(X_a)))
  colSums(X_a/probs) - totals
}

res2 <- multiroot(f = pseudo_ll_calib, start = rep(-1,5))
res2

par <- res2$root
X_a <- samples[samples$nonprob == 1, c("x0", "x1", "x2", "x3", "x4")]
probs <- as.numeric(exp(par %*% t(X_a)) / (1 + exp(par %*% t(X_a))))
colSums(X_a/probs)
weighted.mean(samples[samples$nonprob == 1, "y30"], as.numeric(1/probs))

```

